---
title: "README"
author: "Michael Treadwell"
date: "March 13, 2017"
output: md_document
---

# alterryx

[![Travis-CI Build Status](https://travis-ci.org/mtreadwell/alterryx.svg?branch=master)](https://travis-ci.org/mtreadwell/alterryx)

`alterryx` provides access to each of the Alteryx Gallery API
endpoints. With `alterryx` users can:

* Retrieve information on Alteryx Gallery resouces like apps, workflows, and
macros

* Queue jobs for applications

* Retrieve the status and output of jobs

In order to use this package, you will need to have [a private gallery
API key and secret](https://community.alteryx.com/t5/Alteryx-Knowledge-Base/Private-Gallery-API-Key-and-Secret/ta-p/22009)

# Running Jobs in the Gallery

## Setup

Once you have obtained your API key and secret set them as global options.
Though it is not necessary, it will save you typing later if you also set your
Alteryx Gallery URL as an option.

```{r alterryx_setup}
alteryx_api_key <- "ALTERYX_API_KEY"
alteryx_secret_key <- "ALTERYX_API_SECRET"
alteryx_gallery <- "https://yourgallery.com/gallery"

options(alteryx_api_key = alteryx_api_key)
options(alteryx_secret_key = alteryx_secret_key)
options(alteryx_gallery = alteryx_gallery)
```

## Your Apps
Access to Alteryx Gallery resources like workflows, applications, and macros
are managed through studios. Your account has a subscription id which
determines what you can access. For the purpose of this package, when you see
the term 'resource' that can refer to anything published to the Alteryx Gallery
like workflows, applications, and macros. When you see 'application' or 'app'
it specifically refers to files with the extension _.yxwz_ that are published
to your Gallery.

The resources you can access are obtained using `get_app`.

### Search Apps
```{r get_all_resources, eval = FALSE}
subscription <- get_app()
```

You now have a `list` containing all of the resources you can access. If you
are a power user, this is probably going to be a long list. To pare down the
list use the `request_params` parameter.

If you wanted to see only the five most recently uploaded resources, you can
use the `limit` and `sortField` parameters.

```{r get_some_resources, eval = FALSE}
request_params <- list(
  limit = "5",
  sortField = "uploaddate"
)

subscription <- get_app(request_params)
```

### Non-applications
There is a reason to differentiate between 'resources' and 'apps'. `get_app`
will return all resources that you can access via your subscription that match
the search parameters. However, only 'apps' can be used with the rest of the
API functions. To make sure that `get_app` only returns apps, use
`packageType = "0"` as a request parameter.

```{r get_some_apps, eval = FALSE}
request_params <- list(
  packageType = "0",
  limit = "5",
  sortField = "uploaddate"
)

subscription <- get_app(request_params)
```

### Search Apps by Name
If you are looking for a specific app, it might be easiest to simply search for
it by name. Lets say we are looking for an application named "api_tester.yxwz".

```{r search_app_by_name, eval = FALSE}
request_params <- list(
  packageType = "0",
  search = "api"
)

subscriptions <- get_app(request_params)
app <- subscriptions[[1]]
```

In this case, the app I was looking for was the first result in the list.

### Download a Specific App
If you would like to work with the application in Alteryx, you can use
`doanload_app` to download the application as a _.yxzp_ file.
```{r get_specific_app, eval = FALSE}
download_app(app)
```

## Queueing a Job
Now that I have the app I want, I want to queue a job for it. A 'job' is one
run, a single iteration of an app.

### App Questions
Most of the time, applications have questions that must be answered in order
for the app to run. For example, an app that performs trade area analysis might
ask you to specify a radius for the trade area. These questions are set by the
app author when they create the application in Alteryx Designer. My app,
"api_tester.yxwz" has a single question: "How long should this application
run?"

If you don't have your app questions memorized, use `get_app_questions`.

```{r get_app_questions, eval = FALSE}
questions <- get_app_questions(app)
```

### Build the Answers
Each question has a _name_ and a _value_. In my specific case, the question
name is "runtime" and the default value is "1". Because this app
"api_tester.yxwz" was built specifically to test this API client, the "runtime"
question will simply determine how long the app will run.

I would like the app to run for "3" minutes. Use `build_answers` to format the
answers.

```{r build_answers, eval = FALSE}
name_values <- list(
  name = "runtime",
  value = "3"
)
answers <- build_answers(name_values)
```

If your application has multiple questions, send each as a `list` to
`build_answers`

```{r build_many_answers, eval = FALSE}
name_values1 <- list(
  name = "one",
  value = "1"
)

name_values2 <- list(
  name = "two",
  value = "2"
)

multiple_answers <- build_answers(name_values1,
                                  name_values2)
```

### Queue the Job
Once you have the answers to the app questions you can queue the job.

```{r queue_job, eval = FALSE}
job <- queue_job(app, answers)
```

If the server administrator has granted you access to set job priority, you
can add priority using the `priority` parameter. Valid priority values are
"low", "medium", "high", and "critical".

```{r queue_job_priority, eval = FALSE}
job <- queue_job(app, answers, priority = "high")
```

The `job` will begin with status "Queued". Poll the job using `get_job` to
update the job status.

```{r poll_job, eval = FALSE}
job <- get_job(job)
```

## Job Output
Most Alteryx jobs contain an output tool that writes data once the job is
complete. Use `get_job_output` to retrieve the results as a `data.frame`. The
result will be a list with one element for each valid output. 'Valid' in this
case means an output from an Alteryx output tool that can be converted into a
`data.frame`.

A job needs to have a "Completed" status before outputs can be retrieved.

```{r get_job_output, eval = FALSE}
output <- get_job_output(job)
```

### Invalid Job Outputs
All outputs cannot be properly converted to a `data.frame`. If your job
contains outputs that cannot be converted, `get_job_output` will issue a
warning by default and skip the 'invalid' outputs.

In order to be properly converted, your output must be written in _csv_ or
_yxdb_ format from the Alteryx app published to Gallery.

# Migrating Apps, Macros, and Workflows

## Setup

For migration, you will need keys and access to the admin Gallery API.
Non-admin keys will not work for migration and admin keys will not work for
normal API operation.

Once you have obtained your admin API key and secret set them as global options.
Though it is not necessary, it will save you typing later if you also set your
Alteryx Gallery URL as an option.

```{r alterryx_source_setup, eval = FALSE}
alteryx_api_key <- "ALTERYX_API_KEY"
alteryx_secret_key <- "ALTERYX_API_SECRET"
alteryx_gallery <- "https://yourgallery.com/gallery"

options(alteryx_api_key = alteryx_api_key)
options(alteryx_secret_key = alteryx_secret_key)
options(alteryx_gallery = alteryx_gallery)
```

Migration involves movement from one environment to another, so you will also
need access to the admin API for the target environment.

```{r alterryx_target_setup, eval = FALSE}
target_alteryx_api_key <- 'TARGET_ALTERYX_API_KEY'
target_alteryx_secret_key <- 'TARGET_ALTERYX_API_SECRET'
target_alteryx_gallery <- "https://your-other-gallery.com/gallery"

options(target_alteryx_api_key = target_alteryx_api_key)
options(target_alteryx_secret_key = alteryx_secret_key)
options(target_alteryx_gallery = target_alteryx_gallery)
```

*NOTE* You can migrate from and to the same enviornment.

## Migratable Apps

Users and admins can mark all resources on Alteryx Gallery as "ready for
migration." Once marked, admins can get a list of all resources that are ready.
The function returns all apps in the Gallery that are ready for migration.

```{r migratable_apps, eval = FALSE}
migratable_apps <- get_migratable()

# Searching for resources marked for migration... 
# 2 resources found... 
# 
#    APP_1.yxwz 
#    APP_2.yxwz
```

*Known Issue* `get_migratable` has an option to filter resources by a list of
specific subscriptions (private studios) but the functionality is not working
properly.

```{r migratable_apps_known_issue, eval = FALSE}
get_migratable(
  subscription = list(
    subscriptionIds = "subscription1,subscription2"
  )
)
```

## Downloading and Migrating

Once you've obtained a list of apps to migrate, you can migrate them using
`publish()`. The function downloads the app to the given directory and then
publishes to the Gallery using the credentials and information you set as
options.

When publishing, you can provide a form which will determine certain options
for the app in the target environment. If you do not provide a form, the form
will be generated for you using the app information in the source environment.

* `name`: Give the app a new name
* `owner`: Give the app a new owner (must exist in target system)
* `validate` Perform a validation run prior to publishing?
* `isPublic` Place the app in the Public Gallery?
* `sourceId` See below
* `workerTag` Assigned worker name
* `canDownload` Can others download the app?

If `sourceId` is left blank, it will create a new copy of the published app in
the target enviornment. If you provide the app ID from the _source_
environment, it will overwrite the appropriate app in the _target_ environment.

```{r publish, eval = FALSE}
form <- list(
  name = "APP_1.yxwz",
  owner = "yourname@domain.com",
  validate = "false",
  isPublic = "false",
  sourceId = "",
  workerTag = "",
  canDownload = "false"
)

publish(
  migratable_app,
  migration_directory = "c:/temp",
  form = form
)

# APP_1.yxwz 
# --------------- 
#   Downloading workflow... 
#   Saving workflow... 
#   Publishing workflow to target environment... 
#   Done. Publish in target environment as 2acd7c 
```

## Toggle Migration Flag

Once the app has been published to the target enviornment, you will likely want
to turn off the "Ready to Migrate" flag in the source environment.

```{r toggle_migratable, eval = FALSE}
toggle_migratable(migratable_app)
```

## Automation

I've provided a convenient wrapper to automate migrating workflows:
`migrate()`. This function identifies all resources ready for migration and
publishes them to the target enviornment. If you provide a form for migration,
it will use the information provided but change the app name and ID for each
item marked for migration.

```{r migrate, eval = FALSE}
migrate(
  form = form,
  migration_directory = "c:/temp"
)

# Beginning Migration... 
# 
# Searching for resources marked for migration... 
# 2 resources found... 
# 
#    APP_1.yxwz 
#    APP_2.yxwz 
# 
# APP_1.yxwz 
# --------------- 
#   Downloading workflow... 
#   Saving workflow... 
#   Publishing workflow to target environment... 
#   Done. Publish in target environment as 2acd7c 
# 
# 
# APP_2.yxwz 
# --------------- 
#   Downloading workflow... 
#   Saving workflow... 
#   Publishing workflow to target environment... 
#   Done. Publish in target environment as 2acd97 
# 
# Toggling Migrate Flag... 
# Done
```



